### 2022-10-13 README.Rmd
update

# Interoceptive Accuracy Scale Re-analysed (IAS-R)

@@

- Murphy, J., Brewer, R., Plans, D., Khalsa, S. S., Catmur, C., & Bird, G. (2020). [Testing the independence of self-reported interoceptive accuracy and attention](https://journals.sagepub.com/doi/full/10.1177/1747021819879826). Quarterly Journal of Experimental Psychology, 73(1), 115-133. https://doi.org/10.31234/osf.io/fqgu4

**Aim**. Re-analyze the raw data from [Murphy et al. (2020)](https://journals.sagepub.com/doi/full/10.1177/1747021819879826) (study 1) regarding the factor structure of the Interoceptive Accuracy Scale (IAS) by using the *Method Agreement Procedure* for the number of factors to extract and performing a CFA model comparison.

**Conclusion**. EFA suggested a 1-factor and 4-factors solutions, but the latter was favoured by CFA. Additionally, CFA suggested that a 5-factors model (obtained by separating *Nociception* from *Interoception*) was superior:

- *Interoception*: Heart, Hungry, Breathing, Thirsty, Temperature, Sexual arousal.
- *Nociception*: Muscles, Bruise, Pain.
- *Expulsion*: Burp, Sneeze, Wind.
- *Elimination*: Vomit, Defecate, Urinate.
- *Skin*: Itch, Tickle, Affective touch.

The final revised scale, made of 18 items (6 for interoception and 3 per secondary dimension), is available below.

## Data

Data from the [study 1](https://osf.io/3m5nh/?view_only=a68051df4abe4ecb992f22dc8c17f769) was downloaded from OSF.

@@

report::report_participants(df, age = "Age", sex = NA, gender = "Gender")
```

@@

The *Method Agreement Procedure*

@@

We compared the two models suggested by EFA (1-factor and 4-factors), with two additional model based on the latter: one with 2 items removed (Blood sugar and Taste), and one with additionally the *Interoception* factor split into two (with the pain-related items grouped together).

@@

cfa4mod <- "
Skin =~ Affective_touch + Tickle + Itch
Expulsion =~ Sneeze + Cough + Wind + Burp
Interoception =~ Heart + Hungry + Breathing + Thirsty + Temp + Sex_arousal + Muscles + Bruise + Pain
Elimination =~ Urinate + Defecate + Vomit
" |> 
lavaan::cfa(data=data)

@@

anova(cfa1, cfa4, cfa4mod, cfa5)
```

Additionally, we removed the least loaded items of expulsion (cough) to improve the balance (3 items per secondary scales, and 6 for interoception).

@@

```{r, message=FALSE, warning=FALSE}
parameters::parameters(cfa5) |> 
  display()
```

@@

## Final Scale

>>>>

# Interoceptive Accuracy Scale Revised (IAS-R)

@@

<sub>A reanalysis of Murphy, J., Brewer, R., Plans, D., Khalsa, S. S., Catmur, C., & Bird, G. (2020). [Testing the independence of self-reported interoceptive accuracy and attention](https://journals.sagepub.com/doi/full/10.1177/1747021819879826). Quarterly Journal of Experimental Psychology, 73(1), 115-133. https://doi.org/10.31234/osf.io/fqgu4</sub>

We present a re-analysis of the data from [**Murphy et al. (2020)**](https://journals.sagepub.com/doi/full/10.1177/1747021819879826) (study 1) regarding the factor structure of the **Interoceptive Accuracy Scale (IAS)**. In particular, we apply the *Method Agreement Procedure* to estimate the optimal number of latent factors, and perform a further model comparison using Confirmatory Factor Analysis (CFA).

Exploratory Factor Analysis suggested a 1-factor and 4-factors solutions, but the latter was favoured by CFA. Further comparison suggested that a 5-factors model (obtained by separating *Nociception* from *Interoception*) had a superior fit. The 5 factors are:

- **Interoception**: Heart, Hungry, Breathing, Thirsty, Temperature, Sexual arousal.
- **Nociception**: Muscles, Bruise, Pain.
- **Expulsion**: Burp, Sneeze, Wind.
- **Elimination**: Vomit, Defecate, Urinate.
- **Skin**: Itch, Tickle, Affective touch.

The final revised scale, made of 18 items (6 for interoception and 3 per secondary dimension), is available [below](#IAS-R).

## Participants

@@

paste0(
  "Data from the [study 1](https://osf.io/3m5nh/?view_only=a68051df4abe4ecb992f22dc8c17f769) (Murphy et al., 2020), downloaded from OSF, included ",
  report::report_participants(df, age = "Age", sex = NA, gender = "Gender"),
  "."
)
```

@@

The *Method Agreement Procedure* suggested that 1 factor was optimal, followed by 4 factors.

@@

@@

anova(cfa1, cfa4)
```

The 4-factors model suggested by EFA presented a significantly better fit than the 1-factor model. We then compared that model to two variants: one with 2 items removed (Blood sugar and Taste), and another with, additionally, the *Interoception* factor split into two (with the pain-related items grouped together). The latter model (*CFA-5*), was significantly superior to the others. Additionally, we removed the least loaded items of expulsion (cough) to improve the balance (3 items per secondary scales, and 6 for interoception), which significantly improved the model fit.

```{r, message=FALSE, warning=FALSE}
cfa4mod <- "
Skin =~ Affective_touch + Tickle + Itch
Expulsion =~ Sneeze + Cough + Wind + Burp
Interoception =~ Heart + Hungry + Breathing + Thirsty + Temp + Sex_arousal + Muscles + Bruise + Pain
Elimination =~ Urinate + Defecate + Vomit
" |> 
lavaan::cfa(data=data)

@@

anova(cfa4, cfa4mod, cfa5)
```

@@

@@

```{r, message=FALSE, warning=FALSE}
performance::performance(cfa5mod) |> 
  display()
```

```{r, message=FALSE, warning=FALSE}
parameters::parameters(cfa5mod) |> 
  display()
```

## IAS-R
